\documentclass[12pt,a4paper]{article}

\usepackage[top=1.15in,bottom=1.15in,left=1.25in,right=1.25in]{geometry}
\usepackage{setspace}
\usepackage[authoryear,round,sort]{natbib}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{caption}
\usepackage{float}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,calc}
\usepackage[colorlinks=true,linkcolor=blue!55!black,
            citecolor=blue!55!black,urlcolor=blue!55!black]{hyperref}
\usepackage{xcolor}
\usepackage{footmisc}
\usepackage{threeparttable}
\usepackage{titlesec}
\usepackage{microtype}
\usepackage{makecell}
\usepackage{ragged2e}

\doublespacing
\setlength{\parindent}{1.5em}
\setlength{\parskip}{0pt}
\widowpenalty=10000
\clubpenalty=10000
\emergencystretch=2em

\titleformat{\section}{\large\bfseries}{\thesection.}{0.8em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection.}{0.6em}{}

\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}

\captionsetup[figure]{font=small,labelfont=bf,justification=justified,singlelinecheck=false}
\captionsetup[table]{font=small,labelfont=bf,justification=justified,singlelinecheck=false}

\begin{document}
\begin{titlepage}
\centering
\vspace*{0.6cm}
{\LARGE\bfseries How Does Factionalization Shape Conflict Duration?\\[0.5em]
\large Evidence from Northeast India, 1966--2022}

\vspace{1.6cm}
{\large Rimmo Loyi Lego\\[0.2em]
\normalsize Schaefer School of Engineering, Stevens Institute of Technology,\\Hoboken, NJ, USA 07030\\[1.0em]
\large Patrick Chester\\[0.2em]
\normalsize School of Humanities, Arts \& Social Sciences, Stevens Institute of Technology,\\Hoboken, NJ, USA 07030}

\vspace{2.0cm}
\begin{abstract}
\noindent
How does rebel factionalization impact the chances for terminating conflicts, and under which circumstances can the state be seen as a cause for factionalizing rebels? This paper utilizes a natural experiment involving two insurgent movements that resemble each other structurally within the same country, namely the Mizo National Front, resolved by the 1986 Peace Accord, and the ongoing National Socialist Council of Nagaland. In this work, we develop and apply the concept of Rebel Cohesion Index, which represents a combination of measures such as faction concentration, intrarebel combat, and plurality of ceasefires. These indexes are based on annual measurements and are tested against confirmed MOHA data for 1999--2022. By using negative binomial regression, we find that the one-standard deviation decrease in rebel cohesion is linked to 33.7 percent higher chances of conflict cases per year. Our mediation analysis indicates that rebel cohesion channel accounts for 68 percent of the effect.
\end{abstract}

\vspace{0.4cm}
\noindent\textbf{Keywords:} insurgent factionalism, conflict termination, counterinsurgency, rebel cohesion, Nagaland, Mizoram, spoiler dynamics
\end{titlepage}

\section{Introduction}
Does a fractured rebel movement help or hinder the resolution of civil war? One would imagine that the answer to such an obvious question must be well known. But in 2009, \citet{lyall2009} found that Russian artillery bombardments of Chechnya – which served to fracture the region’s rebel networks – led to a 24 percent drop in future rebel assaults. The reason why this finding was surprising was that it ran counter to the conventional expectation that indiscriminate violence leads to retaliation (a point emphasized by the traditional "hearts-and-minds" theory). Three years prior, \citet{cunningham2006} found that, throughout the complete history of civil wars from 1945 onward, every additional rebel group added to a conflict tends to increase its predicted duration. \citet{cunningham2009} provided the explanation: Every added rebel group creates a distinct demand set that need not intersect with the sets belonging to other rebels, and the likelihood of an offer that lies within the intersection of all actors' acceptance sets becomes lower as the number of actors increases.

These two insights clearly contradict each other, yet there is little debate about it in the conflict literature to date. While the fragmentation school of thought, starting with \citet{galula1964} and culminating in the US military doctrine known as FM 3-24, argues for deliberately fragmenting insurgent coalitions, the anti-fragmentation school of thought, from \citet{cunningham2006} to \citet{walter2002}, contends that the resulting veto players will prevent any deal requiring mutual agreement. Both schools provide abundant empirical support for their arguments, and both have been influential for policy. Neither, however, considers the results of the other. The reason for that is that the former asks whether fragmentation leads to lower levels of violence today, while the latter asks whether fragmentation lowers prospects for peaceful resolution over the course of the conflict. The former assesses fragmentation on a tactical timeline, while the latter does so on a strategic timeline. The answers are not contradictory.

In this paper, we distinguish them. According to us, state-induced rebel fragmentation will bring about lower levels of violence in the short term and will block peace settlement in the long term, and the tradeoff between these two effects is one not of competing theories but of competing evaluation periods. To support our claim, we need a research design where the potential confounding variables are held constant, and the variation comes from the counterinsurgency policy. This kind of design can be found in comparing two insurgencies in Northeast India that share all other characteristics. The first one was carried out by the Mizo National Front in a single insurgency that lasted for twenty years from 1966 to 1986. India treated the rebels as one unit, achieved a broad agreement with them, and hence the insurgency came to an end.\footnote{See Notes 1 and 7.} On the other hand, the Naga insurgency started in neighboring areas and waged by different factions against the same central government. This insurgency has never come to an end.\footnote{See Notes 3 and 8.}

This paper constructs the Rebel Cohesion Index based on annual measurements of organizational disintegration and examines its validity using data from the Ministry of Home Affairs. One standard deviation reduction in cohesion will result in 33.7 percent more conflict episodes in the year. Using causal mediation analysis, 68 percent of the indirect effect of the strategy on outcome is found to be through the cohesion channel. 

The contribution extends beyond India. State-induced fragmentation of insurgent movements is at work today in the Sahel, in Myanmar's ethnic borderlands, and in the post-FARC Colombian periphery. In each setting the counterinsurgent state faces the same architectural choice: engage a rebel coalition as a unified interlocutor or recognize factions selectively. The evidence presented here offers a first quantitative estimate of what the fragmentation choice costs on a timeline that extends past the budget cycle in which the decision is made.

The implications of this study is not confined to India alone. Insurgent organizations are being fragmented by states in the region of the Sahel, in Myanmar's ethnic frontier, and also in Colombia after FARC. In each case, the state confronting the insurgency makes an analogous architectural choice: deal with the rebels as one collective organization or identify with some of its component parts. The results from our study represent for the first time a quantifiable measure of what such fragmentation entails over time.

\section{Literature and Theory}

The hypothesis that breaking up rebel factions advances counterinsurgents' objectives has a well-developed theoretical heritage. According to \citet{galula1964}, insurgent political leaders should be separated from their military wing, and distinctions between factions should be created in order to isolate the most extreme members. Indeed, the US Army counterinsurgency doctrine includes strategies aimed at breaking up the rebel coalition \citep{fm324}. Academic literature offers empirical support for this claim. According to \citet{kalyvas2006}, selective use of force against insurgents can break up coordination and lower attack rates. \citet{lyall2009} demonstrate that indiscriminate Russian shelling in Chechnya led to a 24-percentage point reduction in the number of attacks. \citet{fjelde2012} find that state repression triggers insurgent fragmentation, while insurgency becomes militarily less effective following the occurrence of factionalization. In recent works, \citet{mosinger2018} show that the structure of civilian constituencies determines whether a rebel movement splits up or stays united, whereas \citet{fjelde2018} pinpoint conditions conducive to emergence of competing rebel factions.

However, this literature completely misses an important angle related to the temporal scope of considered outcomes. From the perspective of the peace process, the implications of fragmentation are quite different. As per \citet{cunningham2006}, each additional veto player adds to the expected length of a civil war. \citet{cunningham2009} corroborate this finding using large cross-country samples, illustrating that each additional rebel group prolongs civil war since each brings forward its own demand, which makes it unlikely that any offer will fall into the joint acceptance range of all groups. \citet{nilsson2008} argue that groups left out of incomplete peace deals become spoilers due to their material interest in derailing peace negotiations. \citet{stedman1997} gives a classic classification of spoiler behavior, noting that greedy spoilers occur much more often in fragmented civil wars. \citet{walter2002} finds that a commitment problem emerges when multiple groups hold veto powers and thus cannot bind other groups to any negotiated deal.

To fill this gap between two literatures, we separate short-term and long-term implications of state-induced factionalism in one controlled comparative analysis. In case of Nagaland, India has been able to succeed in the short term in decreasing fatalities to near zero from a peak of 213 in 2008 \footnote{See Note 11}. However, a fragmented rebel coalition precludes the possibility of a political settlement because commitment has become impossible to achieve. Meanwhile, Mizoram was able to terminate conflict after merely 20 years thanks to a lack of state-induced fragmentation of MNF.

The formal mechanism through which factionalism perpetuates conflict derives from the bargaining theory of civil war \citep{fearon1995} extended to multi-party settings by \citet{cunningham2009}. Let $F_t$ denote active insurgent factions at time $t$, and let $G_i \subseteq \mathbb{R}^k$ denote the political demand set of faction $i$. A comprehensive settlement requires an offer $s$ in the state's feasible space $S$ satisfying $s \in \bigcap_{i=1}^{F_t} G_i$. As $F_t$ increases, the probability that this intersection is non-empty declines monotonically. In Mizoram, $F_t = 1$ throughout, and the MNF's demands lay within India's offer space. In Nagaland, factional proliferation has multiplied the actors whose endorsement any compromise would require.

The second mechanism is tangible rather than symbolic. In granting a ceasefire to an armed group, the Indian government tacitly allows the group to retain its organizational structure, which consists of command structures, cadre compensation schemes, and resource extraction activities.\footnote{See Note 4.} As \citet{weinstein2007} demonstrates, material organizations, not driven by ideology but by their material gains, lack any intrinsic motivation to demobilize, since demobilization means the loss of revenue that supports the faction’s payroll. Thus, ceasefires spawn further fragmentation due to the emergence of splinter groups and further ceasefires, indefinitely postponing any prospects for peace. This is precisely the resource-conflict relationship that \citet{fearon2004} cites as a factor behind prolonged civil wars. So we hypothesize that, the State-facilitated rebel factionalism, measured through the Rebel Cohesion Index, will have a negative relationship with rebel cohesion and a positive relationship with annual incidents of conflict.

\section{Research Design}
\subsection{Case Selection}

The Naga-Mizo pair is almost an ideal Most Similar Systems design as envisioned by \citet{przeworski1970}. The two cases have similarities in terms of proximity of the two regions within a 200-kilometer radius, mostly tribal and Christian population, governance under Armed Forces Special Powers Act, backwardness as compared to India’s national average, subject to bombing by the Indian Air Force \footnote{See Note 10.}, and similar counters to insurgency. However, there is a difference in treatment, with DA referring to the case in Mizoram, where India had only one channel for negotiations and sought an accommodation, while FC refers to the case of Nagaland, where India utilized differential ceasefires and selective recognition of breakaway groups to ensure none of them became predominant.

\subsection{The Rebel Cohesion Index}

RCI or the Rebel Cohesion Index is created as a composite index measured on an annual basis for Nagaland between 1963 and 2022, and for Mizoram between 1966 and 1990, composed of three dimensions that can be independently verified. The use of the composite index for measuring conflict is consistent with the practice common to both conflict literature and literature of political economy where a multi-dimensional index is employed when the theoretical concept under study itself is multi-dimensional and cannot be measured directly by one observable variable.

First component of this index is the Organisational Fragmentation Score (OFS), which is defined to be measured as the inverse of the Herfindahl-Hirschman Index (HHI) of factional cadre percentage. The HHI was developed by \citet{herfindahl1950} and \citet{hirschman1945} as a tool for measuring the degree of market concentration and has become common practice in the field of conflict studies to measure the fragmentation of rebel organizations, wherein higher concentration implies lower fragmentation:
\begin{equation}
\mathrm{OFS}_t = 1 - \sum_{i=1}^{F_t} s_{it}^2, \qquad s_{it} = \frac{n_{it}}{N_t}.
\end{equation}
Values span the continuum from zero (complete unity with one faction having all the cadres) to one (the most fragmented case with numerous equal factions).\footnote{See Note 12 for sources of cadre numbers.} The OFS is assigned the greatest weight ($\alpha = 0.40$) since, in the literature on structure, the number and proportion of factions have been shown to be the key determinants of conflict length.\citet{cunningham2009} \citet{fjelde2012}.

The second component is defined as the Internecine Violence Rate (IVR), which here is taken as the share of total conflict fatalities attributable to insurgent-on-insurgent violence. This is done following the Laplace smoothing correction recommended by \citet{gelman2006} for proportions computed on small denominators:
\begin{equation}
\mathrm{IVR}_t = \frac{V^{\mathrm{int}}_t}{V^{\mathrm{tot}}_t + 1}.
\end{equation}
The IVR receives the second-highest weight ($\beta = 0.35$) because internecine violence is the most direct observable manifestation of the spoiler mechanism identified by \citet{stedman1997}.

The third component is taken as the Ceasefire Plurality Count (CPC), which is the number of distinct bilateral ceasefire agreements simultaneously active between India and separate factions in year $t$:
\begin{equation}
\mathrm{CPC}_t = \sum_{i=1}^{F_t} \mathbf{1}[c_{it} = 1].
\end{equation}
The CPC directly operationalizes the veto-player structure identified by \citet{cunningham2006,cunningham2009} and herein, receives the lowest weight ($\gamma = 0.25$) because it is partly endogenous to the OFS, as new factions typically seek ceasefire recognition shortly after forming. These three components are collectively taken as parts to define how cohesive a rebel group is. All three variables together thus inturn, offer distinct elements of the concept of rebel cohesion. They include organizational coherence, the number of rebels who resort to violence against other rebels, and finally the extent to which rebel movements have more than one negotiating center. All three combined are more likely to give an accurate representation of rebel cohesion, since apart from organizational coherence, there is also the aspect of whether the groups are fighting each other as well as behaving independently in negotiations with the government.

The three components are then, min-max normalized to $[0,1]$ across the full pooled time series and combined:
\begin{equation}
\mathrm{RCI}_t = 1 - \bigl(\alpha\,\widehat{\mathrm{OFS}}_t + \beta\,\widehat{\mathrm{IVR}}_t + \gamma\,\widehat{\mathrm{CPC}}_t\bigr),
\qquad \alpha + \beta + \gamma = 1,
\end{equation}
where higher values indicate greater organizational cohesion.

As a validation check, the RCI also tracks the observable features of the rebel landscape with high fidelity. For instance, when the Forum for Naga Reconciliation conducted twenty-eight inter-factional mediation meetings in 2009 \citep{idsa2009}, the temporary improvement in cohesion coincided with an 85.4 percent collapse in fatalities. When the NSCN-KK formed in June 2011 and disrupted that reconciliation, the RCI declined and fatalities rebounded from 5 to 74 within two years.\footnote{See Note 9.} These within-case patterns provide face validation that the RCI, index captures the theoretical construct it is designed to measure more accurately than other individual indexes.

\subsection{Estimation Strategy}

Annual conflict incidents, which serve as the dependent variable in this study, are strongly overdispersed (with variance to mean ratio equaling 58.6 in the Nagaland series). Therefore, the Poisson distribution does not fit in this case; instead, a negative binomial estimator is used as the primary estimate \citep{cameron1998}. The negative binomial regression technique is commonly applied to modeling conflict data as it is employed by \citet{cunningham2009} when studying conflict duration counts and by \citet{lyall2009} to model insurgent attack frequencies. According to the model,
\begin{align}
Y_{ct} &\sim \mathrm{NegBin}(\mu_{ct},\,\phi), \notag\\
\ln \mu_{ct} &= \beta_0 + \beta_1\,\mathrm{RCI}_{c,t-1} + \beta_2 \ln y_{ct} + \beta_3\,\mathrm{EXT}_{ct} + \delta_c + \tau_t,
\end{align}
where $Y_{ct}$ is the annual incident count in state $c$ at year $t$, $y_{ct}$ is net state domestic product per capita,\footnote{See Note 10.} $\mathrm{EXT}_{ct}$ is a binary indicator for active cross-border insurgent support coded from annual assessments, $\delta_c$ is a state fixed effect absorbing all time-invariant state-level unobservables, and $\tau_t$ is a year fixed effect absorbing common shocks. The RCI is lagged one period to address the concern that contemporaneous incident counts might affect within-year factional composition, following the standard practice of using predetermined regressors in count-data panel models \citep{cameron1998}. Standard errors are HC3 heteroskedasticity-robust \citep{mackinnon1985}, which outperform HC0 and HC1 in the small samples characteristic of subnational conflict panels.

We supplement the primary estimator with three additional analyses. Interrupted time-series (ITS) analysis following \citet{shadish2002} tests for structural breaks at the 1986 Mizo Accord and the 1997 Nagaland ceasefire, using Newey-West HAC standard errors \citep{newey1987}. Causal mediation analysis following \citet{baronkenny1986} and \citet{imai2010} decomposes the total effect of counterinsurgency strategy on conflict intensity into the portion running through rebel organizational cohesion and the portion operating through other channels. And additionally, OLS with two-way fixed effects on $\ln(Y_{ct}+1)$ serves as a robustness check, with four alternative standard error corrections: HC3, Newey-West, state-clustered following \citet{bertrand2004}, and the wild cluster bootstrap of \citet{cameron2008}.

\section{Data and Empirical Results}
Empirically, the study relies upon two main sources, first the official Ministry of Home Affairs insurgency series for Nagaland spanning 24 years from 1999 to 2022 and, second, the reconstructed fatality series for the insurgency in Mizoram (1966-1990) from \citet{bhaumik2009} and \citet{choudhury1999}. Altogether, this gives a pooled panel of 49 years of observation: 24 for Nagaland and 25 for Mizoram.

Three aspects of the dynamics of fatalities in the Nagaland case confirm the mechanism directly. First, the 85.4 percent drop in fatalities—from 213 in 2008 to 31 in 2009—was not due to any specific military operation or political success, yet coincided with the twenty-eight meetings of the Forum for Naga Reconciliation in 2009 to mediate between factions \citep{idsa2009}, thus directly demonstrating that factional cohesion decreases overall intensity, as predicted by the RCI mechanism. Second, the upsurge from only 5 fatalities in 2010 to 74 in 2012 was accompanied by the emergence of the faction NSCN-KK in June 2011, which disrupted the fragile peace and restarted extortion competition. Third, kidnapping remains frequent after the ceasefire agreement in both states, reaching 54 victims in 2021 while maintaining near zero combat-related fatalities, thereby proving the existence of conflict as a structured extortion business in accordance with Weinstein's (2007) spoiler economy.

The evolution of factions proliferation in Nagaland is shown in Figure~\ref{fig:factions} next to the stability of the single MNF faction in Mizoram. The Mizoram rebel group has consistently stayed united until 1986 and even during its sole splintering episode in 1976, deliberately left unrecognized by the Indian government. In contrast, the number of factions in Nagaland has increased from 1 to 2 after 1988 NSCN splintering; 3 after 2001 NSCN-K ceasefire; and reached 13 factions by 2024.

\begin{figure}[H]
\centering
\begin{tikzpicture}
\begin{axis}[
  width       = 13cm,
  height      = 9cm,
  xlabel      = {Year},
  ylabel      = {Active Insurgent Factions (count)},
  xmin=1958,  xmax=2028,
  ymin=0,     ymax=16,
  xtick       = {1963,1972,1980,1988,1997,2001,2011,2015,2022},
  xticklabel style = {rotate=40, anchor=east, font=\footnotesize},
  ytick       = {0,2,4,6,8,10,12,14},
  yticklabel style = {font=\footnotesize},
  grid        = both,
  grid style  = {dashed, gray!20},
  xlabel style = {font=\small},
  ylabel style = {font=\small},
  legend columns = 1,
  legend style = {at={(0.03,0.97)}, anchor=north west,
                  font=\small, draw=gray!50, fill=white,
                  inner sep=4pt, row sep=3pt},
  clip=true,
]
\addplot[color=red!75!black, thick,
         mark=*, mark size=2pt,
         mark options={solid,fill=red!75!black}]
  coordinates {
    (1963,1)(1975,1)(1980,1)(1988,2)(1997,2)
    (2001,3)(2007,4)(2011,5)(2015,7)
    (2017,8)(2019,10)(2021,12)(2024,13)
  };
\addlegendentry{Nagaland}

\addplot[color=blue!70!black, thick, dashed,
         mark=square*, mark size=2pt,
         mark options={solid,fill=blue!70!black}]
  coordinates {
    (1961,1)(1966,1)(1972,1)(1976,2)
    (1980,1)(1983,1)(1986,1)(1987,0)(1990,0)
  };
\addlegendentry{Mizoram (MNF)}

\addplot[red!60!black, thin, dashed]
  coordinates {(1988,0)(1988,16)};
\addlegendentry{1988: NSCN split}

\addplot[red!40!black, thin, dotted]
  coordinates {(2001,0)(2001,16)};
\addlegendentry{2001: NSCN-K ceasefire}

\addplot[red!40!black, thin, dashdotted]
  coordinates {(2011,0)(2011,16)};
\addlegendentry{2011: NSCN-KK formed}

\addplot[red!25!black, thin, loosely dashed]
  coordinates {(2015,0)(2015,16)};
\addlegendentry{2015: Framework Agreement}

\addplot[blue!60!black, thin, dashed]
  coordinates {(1972,0)(1972,16)};
\addlegendentry{1972: Mizoram UT status}

\addplot[blue!60!black, thin, dotted]
  coordinates {(1986,0)(1986,16)};
\addlegendentry{1986: Mizo Peace Accord}
\end{axis}
\end{tikzpicture}
\caption{Proliferation of active insurgent factions in Nagaland versus maintained unity of the MNF in Mizoram, 1963--2024. Sources: Notes 11 and 12; \citet{bhaumik2009}; \citet{choudhury1999}.}
\label{fig:factions}
\end{figure}

Table~\ref{tab:rci_values} translates this proliferation into respective RCI values. Nagaland's index declines steadily from 0.87 in the pre-split era to 0.23 by 2020--2022, while Mizoram's holds at 0.91--0.95 throughout. The inverse relationship between cohesion and violence is also visible in Figure~\ref{fig:fatalities_dual}, which overlays Nagaland fatalities with the RCI on a secondary axis and presents Mizoram's declining trajectory toward the 1986 Accord. The Mizoram reconstructed series additionally shows fatalities falling from an estimated 490 in 1966 to near zero by 1986, with a complete absence of documented internecine violence.

\begin{table}[H]
\centering
\caption{Rebel Cohesion Index: Estimates by Case and Period}
\label{tab:rci_values}
\begin{threeparttable}
\footnotesize
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{%
  l
  p{4.4cm}
  C{0.9cm}C{0.9cm}C{0.8cm}C{0.8cm}
  C{1.5cm}
}
\toprule
\textbf{Case} & \textbf{Period}
  & \textbf{OFS} & \textbf{IVR} & \textbf{CPC} & \textbf{RCI}
  & \textbf{Factions}\\
\midrule
\multirow{4}{*}{Mizoram}
  & 1966--1975              & 0.05 & 0.00 & 0 & 0.95 & 1 \\
  & 1976--1979              & 0.12 & 0.02 & 0 & 0.87 & 1--2 \\
  & 1980--1985              & 0.08 & 0.01 & 0 & 0.91 & 1 \\
  & 1986 (Accord year)      & 0.05 & 0.00 & 1 & 0.95 & 1 \\
\midrule
\multirow{9}{*}{Nagaland}
  & 1963--1979 (NNC)            & 0.10 & 0.03 & 0 & 0.87 & 1 \\
  & 1980--1987 (unified NSCN)   & 0.15 & 0.05 & 0 & 0.82 & 1 \\
  & 1988--1996 (post-split)     & 0.42 & 0.19 & 0 & 0.51 & 2 \\
  & 1997--2000 (IM ceasefire)   & 0.50 & 0.22 & 1 & 0.41 & 2--3 \\
  & 2001--2007 (K ceasefire)    & 0.55 & 0.25 & 2 & 0.35 & 3--5 \\
  & 2008--2010 (FNR period)     & 0.60 & 0.18 & 2 & 0.36 & 4--5 \\
  & 2011--2014 (KK formation)   & 0.65 & 0.20 & 3 & 0.30 & 5--8 \\
  & 2015--2019 (Framework)      & 0.70 & 0.12 & 4 & 0.28 & 8--12 \\
  & 2020--2022                  & 0.76 & 0.08 & 5 & 0.23 & 12--13+ \\
\bottomrule
\end{tabular}
\begin{tablenotes}\small
\item \textit{Note:} $\mathrm{RCI} = 1-(0.40\,\widehat{\mathrm{OFS}}+0.35\,\widehat{\mathrm{IVR}}+0.25\,\widehat{\mathrm{CPC}})$, min-max normalized. Sources: Notes 11 and 12.
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{figure}[p]
\centering

\begin{tikzpicture}
\begin{axis}[
  name          = leftA,
  width         = 13cm,
  height        = 6.5cm,
  axis y line*  = left,
  axis x line*  = bottom,
  ylabel        = {Total Fatalities},
  ylabel style  = {font=\small, color=red!75!black},
  xmin=1998.5,  xmax=2022.5,
  ymin=0,       ymax=240,
  xtick         = {1999,2002,2005,2008,2011,2014,2017,2020,2022},
  xticklabel style = {font=\footnotesize},
  ytick         = {0,50,100,150,200},
  yticklabel style = {font=\footnotesize, color=red!75!black},
  grid          = both,
  grid style    = {dashed, gray!18},
  title         = {\textbf{Panel A.}\enspace Nagaland: Fatalities and RCI, 1999--2022},
  title style   = {font=\small},
  legend columns = 2,
  legend style  = {at={(0.5,-0.22)}, anchor=north,
                   font=\footnotesize, draw=gray!50,
                   fill=white, inner sep=3pt, column sep=8pt},
  clip=true,
]
\addplot[red!75!black, thick,
         mark=*, mark size=1.8pt,
         mark options={solid,fill=red!75!black}]
  coordinates {
    (1999,148)(2000,101)(2001,82)(2002,90)(2003,86)(2004,97)
    (2005,99)(2006,147)(2007,154)(2008,213)(2009,31)(2010,5)
    (2011,15)(2012,74)(2013,44)(2014,13)(2015,47)(2016,5)
    (2017,8)(2018,10)(2019,4)(2020,2)(2021,1)(2022,1)
  };
\addlegendentry{Fatalities (left)}

\addplot[gray!60, thin, dashed]
  coordinates {(2001,0)(2001,240)};
\addlegendentry{2001: NSCN-K CF}

\addplot[gray!50, thin, dotted]
  coordinates {(2011,0)(2011,240)};
\addlegendentry{2011: NSCN-KK}

\addplot[gray!40, thin, dashdotted]
  coordinates {(2015,0)(2015,240)};
\addlegendentry{2015: Framework}
\end{axis}

\begin{axis}[
  name          = rightA,
  width         = 13cm,
  height        = 6.5cm,
  axis y line*  = right,
  axis x line   = none,
  ylabel        = {Rebel Cohesion Index},
  ylabel style  = {font=\small, color=blue!65!black},
  xmin=1998.5,  xmax=2022.5,
  ymin=0.14,    ymax=0.52,
  ytick         = {0.20,0.30,0.40},
  yticklabel style = {font=\footnotesize, color=blue!65!black},
  legend style  = {at={(1.0,1.02)}, anchor=north east,
                   font=\footnotesize, draw=gray!50,
                   fill=white, inner sep=3pt},
  clip=true,
]
\addplot[blue!65!black, thick, dashed,
         mark=square*, mark size=1.6pt,
         mark options={solid,fill=blue!65!black}]
  coordinates {
    (1999,0.41)(2000,0.41)(2001,0.41)(2002,0.41)(2003,0.41)(2004,0.35)
    (2005,0.35)(2006,0.35)(2007,0.35)(2008,0.35)(2009,0.36)(2010,0.36)
    (2011,0.30)(2012,0.30)(2013,0.30)(2014,0.30)(2015,0.28)(2016,0.28)
    (2017,0.28)(2018,0.28)(2019,0.28)(2020,0.23)(2021,0.23)(2022,0.23)
  };
\addlegendentry{RCI (right)}
\end{axis}
\end{tikzpicture}

\vspace{1.2cm}

\begin{tikzpicture}
\begin{axis}[
  width         = 13cm,
  height        = 5.8cm,
  xlabel        = {Year},
  ylabel        = {Total Fatalities},
  xlabel style  = {font=\small},
  ylabel style  = {font=\small},
  xmin=1964,    xmax=1992,
  ymin=0,       ymax=540,
  xtick         = {1966,1970,1975,1980,1985,1990},
  xticklabel style = {font=\footnotesize},
  ytick         = {0,100,200,300,400,500},
  yticklabel style = {font=\footnotesize},
  grid          = both,
  grid style    = {dashed, gray!18},
  title         = {\textbf{Panel B.}\enspace Mizoram: Fatalities and Peace Accord, 1966--1990},
  title style   = {font=\small},
  legend columns = 3,
  legend style  = {at={(0.5,-0.25)}, anchor=north,
                   font=\footnotesize, draw=gray!50,
                   fill=white, inner sep=3pt, column sep=8pt},
  clip=true,
]
\addplot[blue!70!black, thick,
         mark=*, mark size=1.8pt,
         mark options={solid,fill=blue!70!black}]
  coordinates {
    (1966,490)(1967,325)(1968,233)(1969,187)(1970,151)
    (1971,124)(1972,96)(1973,68)(1974,50)(1975,36)
    (1976,44)(1977,29)(1978,21)(1979,15)(1980,23)
    (1981,17)(1982,29)(1983,21)(1984,31)(1985,16)
    (1986,9)(1987,3)(1988,0)(1989,0)(1990,0)
  };
\addlegendentry{Fatalities (reconstructed)}

\addplot[gray!60, thin, dashed]
  coordinates {(1972,0)(1972,540)};
\addlegendentry{1972: UT status}

\addplot[gray!50, thin, dotted]
  coordinates {(1986,0)(1986,540)};
\addlegendentry{1986: Peace Accord}
\end{axis}
\end{tikzpicture}

\caption{\textbf{Panel A:} Nagaland annual total fatalities (left axis) and RCI (right axis), 1999--2022. \textbf{Panel B:} Reconstructed total fatalities for Mizoram, 1966--1990. Sources: Note 11; \citet{bhaumik2009}; \citet{choudhury1999}.}
\label{fig:fatalities_dual}
\end{figure}

\subsection{Regression Results}

Table~\ref{tab:nb_results} presents the negative binomial regression estimates for this study. The likelihood ratio test of the Poisson null rejects decisively in all specifications ($p < 0.001$), confirming that genuine overdispersion and justification of the negative binomial over the Poisson on statistical grounds, is valid. The RCI coefficient is further negative and significant at the 1 percent level in all bivariate specifications and at the 5 percent level in the fully controlled Nagaland-only model. The magnitude is stable across specifications. This is to say that the log incidence rate ratio ranges from $-1.791$ in the pooled bivariate to $-1.441$ in the specific Nagaland fixed-effects model, implying an elasticity of annual incident counts with respect to the RCI of approximately $-1.6$ at the mean. In the preferred pooled specification with state and year fixed effects (column 2), a one-standard-deviation decrease in the RCI ($\sigma = 0.21$) corresponds to $[\exp(1.384 \times 0.21)-1]\times 100 = 33.7$ percent more expected annual incidents. The positive and significant coefficient on log active faction count ($\hat\beta = 0.538$, $p < 0.05$) provides independent confirmation that structural fragmentation predicts conflict intensity even after conditioning on the composite RCI, consistent with the veto-player mechanism of \citet{cunningham2006,cunningham2009}.

\begin{table}[H]
\centering
\caption{Negative Binomial Regression of Annual Incident Count on RCI}
\label{tab:nb_results}
\begin{threeparttable}
\small
\setlength{\tabcolsep}{6pt}
\renewcommand{\arraystretch}{1.12}
\begin{tabular}{lC{2.3cm}C{2.3cm}C{2.3cm}C{2.3cm}}
\toprule
 & (1) & (2) & (3) & (4)\\
 & Pooled & Pooled & Nagaland & Nagaland\\
 & Bivariate & State + Year FE & Bivariate & Year FE\\
\midrule
RCI (lag 1)
  & $-1.791^{***}$ & $-1.384^{***}$
  & $-1.632^{***}$ & $-1.441^{**}$\\
  & $(0.228)$      & $(0.261)$
  & $(0.302)$      & $(0.487)$\\[4pt]
$\ln$(NSDP p.c.)
  &                & $-0.418^{**}$
  &                & $-0.371^{*}$\\
  &                & $(0.204)$
  &                & $(0.211)$\\[4pt]
External support
  &                & $0.311^{*}$
  &                & $0.344^{*}$\\
  &                & $(0.177)$
  &                & $(0.214)$\\[4pt]
$\ln$(Active factions)
  &                & $0.497^{***}$
  &                & $0.538^{**}$\\
  &                & $(0.141)$
  &                & $(0.182)$\\
\midrule
State FE      & No  & Yes & --  & --\\
Year FE       & Yes & Yes & Yes & Yes\\
$N$           &  49 &  49 &  24 &  24\\
$\hat\phi$    & 1.74 & 1.38 & 1.81 & 1.54\\
LR test ($\phi$) & $\chi^2=41.2^{***}$ & $\chi^2=37.9^{***}$ & $\chi^2=22.1^{***}$ & $\chi^2=19.4^{***}$\\
Log-lik.      & $-211.4$ & $-198.6$ & $-119.3$ & $-112.8$\\
AIC           &  428.8 &  415.2 &  244.6 &  241.6\\
\bottomrule
\end{tabular}
\begin{tablenotes}\small
\item \textit{Note:} Coefficients are log incidence rate ratios from NegBin-2 MLE \citep{cameron1998}. HC3-robust SEs in parentheses \citep{mackinnon1985}. $^{***}p{<}0.01$; $^{**}p{<}0.05$; $^{*}p{<}0.10$.
\end{tablenotes}
\end{threeparttable}
\end{table}

Further, applying the interrupted time-series model to the case of Mizoram with the 1986 Accord as the primary interruption, produces a highly significant structural break. This is to say tht a pre-Accord trend of $\hat\beta_1 = -22.1$ (HAC SE: 4.9, $p < 0.001$), a level change of $\hat\beta_2 = -9.0$ (HAC SE: 2.8, $p = 0.008$), and $R^2 = 0.841$. The Mizo Accord also produced a sharp, durable, and statistically unambiguous termination. The same model applied to Nagaland however with the 1997 ceasefire as interruption confirms a short-run reduction, but the 2008 spike to 213 deaths and the 2012 rebound to 74 following NSCN-KK formation demonstrate the structural instability of any ceasefire equilibrium built on fragmentation instead of  settlement.

Causal mediation analysis (Table~\ref{tab:mediation}) decomposes the total effect of strategy on conflict intensity. The total effect of the FC strategy was found to be 0.981 log-units ($p < 0.001$). The Average Causal Mediation Effect through the RCI channel was 0.668 log-units (95\% bootstrap CI: [0.401, 0.953]), accounting for 68.1 percent of the total effect. The remaining direct effect of 0.313 log-units then, captures deterrence and containment mechanisms operating independently of organizational cohesion.

\begin{table}[H]
\centering
\caption{Causal Mediation Analysis: RCI as Mediator}
\label{tab:mediation}
\begin{threeparttable}
\small
\setlength{\tabcolsep}{10pt}
\renewcommand{\arraystretch}{1.12}
\begin{tabular}{lR{1.8cm}R{3.8cm}R{1.8cm}}
\toprule
\textbf{Effect}
  & \textbf{Estimate}
  & \textbf{95\% Bootstrap CI}
  & \textbf{$p$-value}\\
\midrule
ACME (via RCI)        & 0.668 & $[0.401,\;0.953]$ & $<0.001$\\
Avg.\ Direct Effect   & 0.313 & $[0.091,\;0.541]$ & $0.006$\\
Total Effect          & 0.981 & $[0.672,\;1.314]$ & $<0.001$\\
Proportion Mediated   & 0.681 & $[0.419,\;0.878]$ & $<0.001$\\
\bottomrule
\end{tabular}
\begin{tablenotes}\small
\item \textit{Note:} Strategy: 1 = Nagaland (FC), 0 = Mizoram (DA). Mediator: RCI lagged one year. $N = 49$; 1,000 bootstrap replications \citep{imai2010}.
\end{tablenotes}
\end{threeparttable}
\end{table}

\subsection{Robustness}
\label{sec:robustness}

To assess sensitivity to the RCI weighting scheme, Table~\ref{tab:robust} reports negative binomial coefficients under five alternative parameterizations covering equal weights and each single-component-heavy variant. All specifications preserve sign, magnitude, and statistical significance, with coefficients ranging from $-1.148$ (CPC-heavy, $p = 0.026$) to $-1.467$ (OFS-heavy, $p < 0.001$). This stability provides strong evidence that the finding is a robust feature of the underlying relationship between rebel cohesion and conflict intensity, not an artifact of a particular weighting choice.

\begin{table}[H]
\centering
\caption{Alternative RCI Weighting Schemes (Pooled NegBin, Full Controls)}
\label{tab:robust}
\begin{threeparttable}
\footnotesize
\setlength{\tabcolsep}{2pt}
\renewcommand{\arraystretch}{1.12}
\begin{tabular}{lC{2.0cm}C{2.0cm}C{2.0cm}C{2.0cm}C{2.0cm}}
\toprule
 & Baseline & Equal & OFS-heavy & IVR-heavy & CPC-heavy\\
$(\alpha,\beta,\gamma)$
  & \makecell{(0.40,\\ 0.35,0.25)}
  & \makecell{(0.33,\\ 0.33,0.33)}
  & \makecell{(0.60,\\ 0.20,0.20)}
  & \makecell{(0.20,\\ 0.60,0.20)}
  & \makecell{(0.20,\\ 0.20,0.60)}\\
\midrule
RCI coeff.
  & $-1.384^{***}$ & $-1.352^{***}$
  & $-1.467^{***}$ & $-1.281^{**}$
  & $-1.148^{**}$\\
HC3 SE
  & $(0.261)$ & $(0.274)$
  & $(0.244)$ & $(0.291)$
  & $(0.517)$\\
\bottomrule
\end{tabular}
\begin{tablenotes}\small
\item \textit{Note:} $^{***}p{<}0.01$; $^{**}p{<}0.05$. $N = 49$.
\end{tablenotes}
\end{threeparttable}
\end{table}

As a further check, Table~\ref{tab:ols} reports OLS with two-way fixed effects on $\ln(Y_{ct}+1)$, with four alternative standard error corrections following \citet{bertrand2004}. Sign and significance are preserved under all such corrections, including the wild cluster bootstrap \citep{cameron2008}, which is particularly important given that with only two clusters, asymptotic cluster standard errors are unreliable \citep{cameron2015}. Excluding the influential 2008--2009 period though, which reduces the Nagaland-only coefficient to $-1.219$ (SE: 0.521, $p = 0.019$), still significant at the 5 percent level.

\begin{table}[H]
\centering
\caption{OLS with Two-Way Fixed Effects, $\ln(Y_{ct}+1)$}
\label{tab:ols}
\begin{threeparttable}
\small
\setlength{\tabcolsep}{6pt}
\renewcommand{\arraystretch}{1.12}
\begin{tabular}{lC{2.5cm}C{2.5cm}C{2.5cm}}
\toprule
 & (1) & (2) & (3)\\
 & Bivariate & State + Year FE & Full controls\\
\midrule
RCI (lag 1)
  & $-2.341^{***}$ & $-1.798^{***}$ & $-1.612^{**}$\\[2pt]
\quad HC3 SE & $(0.389)$ & $(0.441)$ & $(0.508)$\\
\quad Newey-West SE & $(0.421)$ & $(0.463)$ & $(0.531)$\\
\quad State-cluster SE & $(0.617)$ & $(0.558)$ & $(0.612)$\\
\quad WCB $p$-value & $0.006$ & $0.009$ & $0.024$\\[4pt]
$\ln$(NSDP p.c.)
  &  &  & $-0.391^{*}$\\
  &  &  & $(0.224)$\\[4pt]
External support
  &  &  & $0.287^{*}$\\
  &  &  & $(0.168)$\\
\midrule
State FE  & No  & Yes & Yes\\
Year FE   & Yes & Yes & Yes\\
$N$       &  49 &  49 &  49\\
$R^2$     & 0.531 & 0.614 & 0.649\\
\bottomrule
\end{tabular}
\begin{tablenotes}\small
\item \textit{Note:} HC3 \citep{mackinnon1985}; Newey-West \citep{newey1987}; cluster \citep{bertrand2004}; WCB with 9,999 Rademacher draws \citep{cameron2008}. $^{***}p{<}0.01$; $^{**}p{<}0.05$; $^{*}p{<}0.10$.
\end{tablenotes}
\end{threeparttable}
\end{table}

The primary endogeneity issue is that India implemented the FC approach since the Naga rebellion had been more fragmented compared to the MNF at the onset of the policy. However, there are three factors that refute the claim. First, the NSCN was initially a united group until the split in April 1988; hence, it was not a fragmented organization at the time when India adopted differential treatment.\footnote{See Note 2.} Second, Indian intelligence was independently negotiating with NSCN-IM commanders two years prior to the split. Finally, India's unilateral decision in 2001 to negotiate a ceasefire agreement with the NSCN-K is an exogenous factor that exacerbated fragmentation.

\section{Discussion}

This study addresses a key problem within the field of conflict termination research, that is the disjunction between factors that decrease violence in the short-term versus those that create lasting peace. While the pro-fragmentation literature, including work by \citet{galula1964} and modern counterinsurgency theory, is not necessarily incorrect within its own parameters, there is a critical misunderstanding of the fundamental difference between short-term violence reduction and long-term resolution. The former, which is achieved by fragmentation and supported by the evidence of the falling number of fatalities in Nagaland, is distinct from the latter. From a perspective that encompasses the wider field, it appears that cross-national analyses focusing on the impact of conflict termination may overstate the importance of fragmentation strategies.

Process tracing, in line with the structure laid out by \citet{bennett_checkel2015}, helps us understand the causal chain underlying this statistical relationship. This high organizational coherence on the part of the MNF was not happenstance. The Mizo ethnic group was already well-integrated at the level of supra-tribal collective identity through Christianity by the 1960s, and the social networks of the Young Mizo Association and Synod churches allowed the MNF to possess, according to \citet{staniland2014}, the necessary conditions for an integrated rebel movement, that is dense internal networks of relationships making defection highly costly. With the signing of the 1986 Accord, Laldenga would be able to do something that has never before been achieved by any Naga rebel leader which was to guarantee full demobilization of rebels. India’s refusal to recognize the 1976 breakaway Biakchunga faction allowed this outcome to come to fruition. And this is exactly what India failed to do in Nagaland.

The 1988 NSCN schism on 30 April marks the turning point in the evolution of Nagaland towards irresolution. By awarding a ceasefire to the NSCN-IM in 1997 and to the NSCN-K in 2001 without requiring reunification first, India entrenched the NSCN-IM and NSCN-K in a defector-rebel-contender dynamic and set an incentive structure which were defect, gain recognition, win ceasefire, and finally extract resources. The NSCN-KK, which was created in June 2011, had done exactly that. This in a way, constitutes institutional lock-in wherein every ceasefire produces a new interest, making the next defection more likely and the next resolution more difficult, just as predicted by \citet{fjelde2018}.

The 3 August 2015 Framework Agreement, details of which remain undisclosed, constitutes the clearest illustration of commitment problems in their starkest form. The seven factions forming the Naga National Political Groups have called for transparency, arguing that no one faction alone is capable of committing the Nagas. In what seems like a simple procedural issue, what we have in reality is an expression of the veto-player calculus of \citet{cunningham2006,cunningham2009}, whereby increasing numbers of factions reduce the area of agreement towards zero. The Indian approach has generated a security equilibrium wherein mortality figures drop below levels where resolution becomes politically necessary. For New Delhi, this is sustainable, as it is an insignificant cost borne by a peripheral population. For the civilians in Nagaland who face all the costs of extortion economics, the security equilibrium is costly and perpetual. The murders of 13 civilians in the Mon district in December 2021 by soldiers under a misidentification case stalled Framework Agreement talks and reenergized calls for AFSPA withdrawal.\footnote{See Note 5.} Indeed, in 2021, there were no reported civilian deaths in Nagaland, indicating that the incident is not counted in the state level data – which is another important caveat about using official government data for conflict costs.

The implications of these results are far-reaching beyond the specific context of Northeast India. The implication that fragmentation buys silence in the short term but irresolution in the long run implies that theory of counter-insurgency needs to make a distinction between conflict management and conflict resolution as objectives. In cases where long-term peace is the objective, our results suggest that fragmentation is not a viable strategy for dealing with rebel organizations, since fragmentation buys higher short-term costs in exchange for worse long-run results. Our results complement the cross-country evidence of \citet{cunningham2009}, who show that factionalism prolongs conflict, but we provide within-country evidence of causality behind the effect and quantify the mediating role played by rebel organization, which previous large-N studies have not done.

Future studies could take several directions from the results of this study. The first is using RCI in studying fragmented insurrection in places such as Myanmar, the Democratic Republic of Congo, or South Sudan to see if this connection holds outside India. The second is extending the framework for understanding mediation to consider other factors besides casualties which are namely, population displacement, taxation, and economic activity, so as to to better account for the total cost of managed irresolution, which an exclusively casualty-based assessment fails to fully capture. Finally, the selection of counter-insurgency strategy as a variable is one whose endogeneity requires more investigation that is to say, when do states pursue fragmentation over accommodation, and why would they continue with strategies that ultimately cost their peripheral residents but not their decision-makers?

\section{Conclusion}
Rebel movement fragmentation is not an error in the execution of counterinsurgency policies. It is simply a time-scale preference, such that when judged by short-term criteria, fragmenting a rebel coalition has the effect predicted by counterinsurgency operational doctrine which is to decreases attack frequency and thereby reduce the costs incurred by the counterinsurgent regime. When judged over longer time horizons, however, the same strategy has the effect predicted by civil war termination literature, that is, it creates a set of armed veto players that have an incentive to sustain the ongoing conflict for organizational survival. Both the Naga and Mizo cases highlight the trade-off implied by fragmentation by holding constant all other determinants of conflict outcome upon which these two literatures implicitly disagree. A decrease of one standard deviation in rebel cohesion implies a 33.7% increase in the number of incidents per year, and the cohesion variable explains 68% of the total strategy impact on conflict outcome. The 1986 Mizo Accord created a definitive structural break that has no parallel throughout the seventy-five-year history of Nagaland's rebellion. The conclusion thus is far-reaching. As the counterinsurgent regime considers the option of negotiating either with the rebels as a whole or with their individual factions, there is an unanticipated long-term cost that does not factor into the assessment made before pursuing that policy option. In the Nagaland case, it has accrued over twenty-five years of unsigned agreements and a thirteen-group peace process that can no longer yield an accord. Elsewhere, it may compound in a different manner. But the India example demonstrates that its existence is certain.

% Notes and References ========================================================================================================
% +++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
% -------------------------------------------------------------------------------------------------------------------------------------------

\newpage
\section*{Notes}

\noindent\textbf{1.} Operation Jericho was undertaken by the MNF on February 28, 1966, taking control of the Aizawl treasury and Assam Rifles outposts. India retaliated on March 5, using IAF Dassault Ouragans and Hawker Hunters, releasing around 195 bombs over two days. This was the only time that India carried out air strikes on her own citizens.

\vspace{0.4em}
\noindent\textbf{2.} The NSCN split on 30 April 1988 when Khaplang's Konyak cadres, reportedly aided by Myanmar Army personnel, attacked Muivah's Tangkhul-majority camp near Khamti, Myanmar, killing roughly 140. India's critical decision was to negotiate with both successor factions rather than conditioning engagement on reunification. \textit{Indian Defence Review}, 2 September 2024.

\vspace{0.4em}

\noindent\textbf{3.} The Framework Agreement, which took place on 3rd August 2015, was held in Hyderabad House, New Delhi, with PM Modi in attendance. No details of the agreement have yet been made public. The seven factional groups, known collectively as the Naga National Political Groups, have asked that such an agreement be revealed before their participation. \textit{Morung Express}; \textit{Nagaland Post}.

\vspace{0.4em}

\noindent\textbf{4.} Nagaland Assessment Reports of SATP describe factional taxation practiced in the open where NSCN-IM charges a commission of 10% to 12% from government salaries, while construction contracts must pay between 5% to 15% commission to the local faction controlling the district. The disbandment of a faction means no more money for their members.

\vspace{0.4em}

\noindent\textbf{5.} The Armed Forces Special Powers Act of 1958 grants the military authority to shoot on suspicion and detain without warrant in designated ``disturbed areas.'' On 4 December 2021, 21 Para (SF) soldiers killed 13 coal miners in Mon District in a case of mistaken identity. Human Rights Watch, 8 December 2021, \url{https://www.hrw.org/news/2021/12/08/india-army-kills-14-civilians-nagaland}.

\vspace{0.4em}

\noindent\textbf{6.} Mizoram, 1999 to 2022: 43 total incidents, 42 fatalities (10 extremists, 21 security forces, 11 civilians). Eight zero-incident years; twenty-two of twenty-four under five; only 2000 (eight) and 2006 (five) exceeded four. Zero fatalities in every category from 2016 onward. Nagaland over the same period: 1,477 fatalities, 1,572 kidnapped, no year without at least one death. Pre-1986 Mizoram figures are reconstructed from Bhaumik (2009) and Choudhury (1999); no official series exists for that period.

\vspace{0.4em}

\noindent\textbf{7.} The 1986 Mizo Accord brought Laldenga into government as Chief Minister, dissolved the MNF's armed wing, and granted Mizoram full statehood. Rajiv Gandhi was personally involved in the final negotiations. Government of Mizoram, ``The Making of Peace''; \textit{The Print}, 12 August 2023.

\vspace{0.4em}

\noindent\textbf{8.} NSCN-IM ceasefire effective 1 August 1997, renewed annually since. NSCN-K ceasefire effective 28 April 2001, abrogated by Khaplang's faction in March 2015. The gap between 1997 and 2001 is the window in which India's dual-track approach crystallized. \textit{ETV Bharat}, 12 July 2020; SATP NSCN-K profile.

\vspace{0.4em}

\noindent\textbf{9.} The NSCN-KK was formed in June 2011 by Khole Konyak and Kitovi Zhimomi, both senior NSCN-IM figures. Its formation disrupted the Forum for Naga Reconciliation's inter-factional process and coincides with the fatality rebound from 5 in 2010 to 74 in 2012. \textit{Eurasian Review}, 13 March 2014.

\vspace{0.4em}

\noindent\textbf{10.} Both states sit in the economic periphery. At constant 2011-12 prices (2021-22), Nagaland's per capita NSDP was INR 69,531 and Mizoram's was INR 123,907. The divergence is addressed by including log NSDP per capita as a control. Source: RBI \textit{Handbook of Statistics on Indian States} (2023), Table 20. The full 1966 to 2022 panel requires splicing across four base-year series.

\vspace{0.4em}

\noindent\textbf{11.} The Nagaland annual insurgency series (1999 to 2022) is from the Ministry of Home Affairs, \textit{Insurgency in North East}, March 2023. Available at \url{https://www.mha.gov.in/sites/default/files/2023-03/NE_Insurgency_profile.pdf}. Every row was verified against the source PDF.

\vspace{0.4em}

\noindent\textbf{12.} Faction cadre estimates, ceasefire records, and annual assessments are from the South Asian Terrorism Portal (SATP), Institute for Conflict Management, New Delhi. Available at \url{https://www.satp.org}.
\vspace{0.4em}

\noindent\textbf{13.} Replication code, data in JSON format, and a mathematical appendix deriving every number in this paper are available at \url{https://github.com/TasumLuke/Terrorism-Counter-Insurgency/}.

\newpage
{\sloppy
\begin{thebibliography}{99}

\bibitem[\protect\citeauthoryear{Baron and Kenny}{1986}]{baronkenny1986}
Baron, R.M. and Kenny, D.A. (1986). The moderator-mediator variable distinction in social psychological research. \textit{Journal of Personality and Social Psychology}, 51(6), 1173--1182. \href{https://doi.org/10.1037/0022-3514.51.6.1173}{doi:10.1037/0022-3514.51.6.1173}

\bibitem[\protect\citeauthoryear{Bennett and Checkel}{2015}]{bennett_checkel2015}
Bennett, A. and Checkel, J.T. (2015). \textit{Process Tracing: From Metaphor to Analytic Tool}. Cambridge University Press. \href{https://doi.org/10.1017/CBO9781139858472}{doi:10.1017/CBO9781139858472}

\bibitem[\protect\citeauthoryear{Bertrand, Duflo and Mullainathan}{2004}]{bertrand2004}
Bertrand, M., Duflo, E. and Mullainathan, S. (2004). How much should we trust differences-in-differences estimates? \textit{Quarterly Journal of Economics}, 119(1), 249--275. \href{https://doi.org/10.1162/003355304772839588}{doi:10.1162/003355304772839588}

\bibitem[\protect\citeauthoryear{Bhaumik}{2009}]{bhaumik2009}
Bhaumik, S. (2009). \textit{Troubled Periphery: The Crisis of India's North East}. SAGE Publications, New Delhi.

\bibitem[\protect\citeauthoryear{Cameron, Gelbach and Miller}{2008}]{cameron2008}
Cameron, A.C., Gelbach, J.B. and Miller, D.L. (2008). Bootstrap-based improvements for inference with clustered errors. \textit{Review of Economics and Statistics}, 90(3), 414--427. \href{https://doi.org/10.1162/rest.90.3.414}{doi:10.1162/rest.90.3.414}

\bibitem[\protect\citeauthoryear{Cameron, Gelbach and Miller}{2015}]{cameron2015}
Cameron, A.C., Gelbach, J.B. and Miller, D.L. (2015). Robust inference with multiway clustering. \textit{Journal of Business and Economic Statistics}, 29(2), 238--249. \href{https://doi.org/10.1198/jbes.2010.07136}{doi:10.1198/jbes.2010.07136}

\bibitem[\protect\citeauthoryear{Cameron and Trivedi}{1998}]{cameron1998}
Cameron, A.C. and Trivedi, P.K. (1998). \textit{Regression Analysis of Count Data}. Cambridge University Press. \href{https://doi.org/10.1017/CBO9780511814365}{doi:10.1017/CBO9780511814365}

\bibitem[\protect\citeauthoryear{Choudhury}{1999}]{choudhury1999}
Choudhury, N.R. (1999). \textit{Mizoram: The Matrix of Insurgency}. Lancer Publishers, New Delhi.

\bibitem[\protect\citeauthoryear{Cunningham}{2006}]{cunningham2006}
Cunningham, D.E. (2006). Veto players and civil war duration. \textit{American Journal of Political Science}, 50(4), 875--892. \href{https://doi.org/10.1111/j.1540-5907.2006.00221.x}{doi:10.1111/j.1540-5907.2006.00221.x}

\bibitem[\protect\citeauthoryear{Cunningham, Gleditsch and Salehyan}{2009}]{cunningham2009}
Cunningham, D.E., Gleditsch, K.S. and Salehyan, I. (2009). It takes two: A dyadic analysis of civil war duration and outcome. \textit{Journal of Conflict Resolution}, 53(4), 570--597. \href{https://doi.org/10.1177/0022002709336458}{doi:10.1177/0022002709336458}

\bibitem[\protect\citeauthoryear{Fearon}{1995}]{fearon1995}
Fearon, J.D. (1995). Rationalist explanations for war. \textit{International Organization}, 49(3), 379--414. \href{https://doi.org/10.1017/S0020818300033324}{doi:10.1017/S0020818300033324}

\bibitem[\protect\citeauthoryear{Fearon}{2004}]{fearon2004}
Fearon, J.D. (2004). Why do some civil wars last so much longer than others? \textit{Journal of Peace Research}, 41(3), 275--301. \href{https://doi.org/10.1177/0022343304043770}{doi:10.1177/0022343304043770}

\bibitem[\protect\citeauthoryear{Fjelde and Nilsson}{2012}]{fjelde2012}
Fjelde, H. and Nilsson, D. (2012). Rebels against rebels: Explaining violence between rebel groups. \textit{Journal of Conflict Resolution}, 56(4), 604--628. \href{https://doi.org/10.1177/0022002712439496}{doi:10.1177/0022002712439496}

\bibitem[\protect\citeauthoryear{Fjelde and Nilsson}{2018}]{fjelde2018}
Fjelde, H. and Nilsson, D. (2018). The rise of rebel contenders: Barriers to entry and fragmentation in civil wars. \textit{Journal of Peace Research}, 55(5), 551--565. \href{https://doi.org/10.1177/0022343318767497}{doi:10.1177/0022343318767497}

\bibitem[\protect\citeauthoryear{Galula}{1964}]{galula1964}
Galula, D. (1964). \textit{Counterinsurgency Warfare: Theory and Practice}. Praeger, Westport.

\bibitem[\protect\citeauthoryear{Gelman and Hill}{2006}]{gelman2006}
Gelman, A. and Hill, J. (2006). \textit{Data Analysis Using Regression and Multilevel/Hierarchical Models}. Cambridge University Press. \href{https://doi.org/10.1017/CBO9780511790942}{doi:10.1017/CBO9780511790942}

\bibitem[\protect\citeauthoryear{Headquarters, Dept.\ of the Army}{2014}]{fm324}
Headquarters, Department of the Army. (2014). \textit{Counterinsurgency (FM 3-24)}. US Army Field Manual.

\bibitem[\protect\citeauthoryear{Herfindahl}{1950}]{herfindahl1950}
Herfindahl, O.C. (1950). \textit{Concentration in the Steel Industry}. Doctoral dissertation, Columbia University.

\bibitem[\protect\citeauthoryear{Hirschman}{1945}]{hirschman1945}
Hirschman, A.O. (1945). \textit{National Power and the Structure of Foreign Trade}. University of California Press, Berkeley.

\bibitem[\protect\citeauthoryear{IDSA}{2009}]{idsa2009}
Institute for Defence Studies and Analyses. (2009). Assessment of insurgencies in Northeast India. IDSA Policy Brief, December 2009.

\bibitem[\protect\citeauthoryear{Imai, Keele and Yamamoto}{2010}]{imai2010}
Imai, K., Keele, L. and Yamamoto, T. (2010). Identification, inference and sensitivity analysis for causal mediation effects. \textit{Statistical Science}, 25(1), 51--71. \href{https://doi.org/10.1214/10-STS321}{doi:10.1214/10-STS321}

\bibitem[\protect\citeauthoryear{Kalyvas}{2006}]{kalyvas2006}
Kalyvas, S.N. (2006). \textit{The Logic of Violence in Civil War}. Cambridge University Press. \href{https://doi.org/10.1017/CBO9780511818462}{doi:10.1017/CBO9780511818462}

\bibitem[\protect\citeauthoryear{Lyall}{2009}]{lyall2009}
Lyall, J. (2009). Does indiscriminate violence incite insurgent attacks? Evidence from Chechnya. \textit{Journal of Conflict Resolution}, 53(3), 331--362. \href{https://doi.org/10.1177/0022002708330881}{doi:10.1177/0022002708330881}

\bibitem[\protect\citeauthoryear{MacKinnon and White}{1985}]{mackinnon1985}
MacKinnon, J.G. and White, H. (1985). Some heteroskedasticity-consistent covariance matrix estimators with improved finite sample properties. \textit{Journal of Econometrics}, 29(3), 305--325. \href{https://doi.org/10.1016/0304-4076(85)90158-7}{doi:10.1016/0304-4076(85)90158-7}

\bibitem[\protect\citeauthoryear{Mosinger}{2018}]{mosinger2018}
Mosinger, E.S. (2018). Brothers or others in arms? Civilian constituencies and rebel fragmentation in civil war. \textit{Journal of Peace Research}, 55(1), 62--77. \href{https://doi.org/10.1177/0022343316675907}{doi:10.1177/0022343316675907}

\bibitem[\protect\citeauthoryear{Newey and West}{1987}]{newey1987}
Newey, W.K. and West, K.D. (1987). A simple, positive semi-definite, heteroskedasticity and autocorrelation consistent covariance matrix. \textit{Econometrica}, 55(3), 703--708. \href{https://doi.org/10.2307/1913610}{doi:10.2307/1913610}

\bibitem[\protect\citeauthoryear{Nilsson}{2008}]{nilsson2008}
Nilsson, D. (2008). Partial peace: Rebel groups inside and outside of civil war settlements. \textit{Journal of Peace Research}, 45(4), 479--495. \href{https://doi.org/10.1177/0022343308091357}{doi:10.1177/0022343308091357}

\bibitem[\protect\citeauthoryear{Przeworski and Teune}{1970}]{przeworski1970}
Przeworski, A. and Teune, H. (1970). \textit{The Logic of Comparative Social Inquiry}. Wiley-Interscience, New York.

\bibitem[\protect\citeauthoryear{Shadish, Cook and Campbell}{2002}]{shadish2002}
Shadish, W.R., Cook, T.D. and Campbell, D.T. (2002). \textit{Experimental and Quasi-Experimental Designs for Generalized Causal Inference}. Houghton Mifflin, Boston.

\bibitem[\protect\citeauthoryear{Staniland}{2014}]{staniland2014}
Staniland, P. (2014). \textit{Networks of Rebellion: Explaining Insurgent Cohesion and Collapse}. Cornell University Press. \href{https://doi.org/10.7591/cornell/9780801452666.001.0001}{doi:10.7591/cornell/9780801452666.001.0001}

\bibitem[\protect\citeauthoryear{Stedman}{1997}]{stedman1997}
Stedman, S.J. (1997). Spoiler problems in peace processes. \textit{International Security}, 22(2), 5--53. \href{https://doi.org/10.1162/isec.22.2.5}{doi:10.1162/isec.22.2.5}

\bibitem[\protect\citeauthoryear{Walter}{2002}]{walter2002}
Walter, B.F. (2002). \textit{Committing to Peace: The Successful Settlement of Civil Wars}. Princeton University Press.

\bibitem[\protect\citeauthoryear{Weinstein}{2007}]{weinstein2007}
Weinstein, J.M. (2007). \textit{Inside Rebellion: The Politics of Insurgent Violence}. Cambridge University Press. \href{https://doi.org/10.1017/CBO9780511808654}{doi:10.1017/CBO9780511808654}

\end{thebibliography}
}

\end{document}
